---
title: "Homework 6"
author: "Ishita Dutta"
date: "11/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1a)
```{r}
eyec = read.table("eyecontact.txt")
n = 5
a = 2
b = 2

A1 <- subset(eyec, eyec[,2] == 1)
A2 <- subset(eyec, eyec[,2] == 2)
B1 <- subset(eyec, eyec[,3] == 1)
B2 <- subset(eyec, eyec[,3] == 2)

A1B1 <- subset(A1, A1[,3] == 1)
A1B2 <- subset(A1, A1[,3] == 2)
A2B1 <- subset(A2, A2[,3] == 1)
A2B2 <- subset(A2, A2[,3] == 2)

Y1.. = mean(A1[,1])
Y2.. = mean(A2[,1])
Y.1. = mean(B1[,1])
Y.2. = mean(B2[,1])

Y11. = mean(A1B1[,1])
Y12. = mean(A1B2[,1])
Y21. = mean(A2B1[,1])
Y22. = mean(A2B2[,1])

SSE = sum((A1B1[,1]-Y11.)^2)+sum((A1B2[,1]-Y12.)^2)+sum((A2B1[,1]-Y21.)^2)+sum((A2B2[,1]-Y22.)^2)
MSE = SSE/((n-1)*a*b)
SY = sqrt(MSE / (b * n))

ME1a =qt(1 - 0.005, (n-1)*a*b) * SY
cat("Confidence interval for Y1.. is [", Y1.. - ME1a, "," , Y1.. + ME1a,"]." )

```
The interpretation of this is that I am 99% confident that the true mean of the treatments where there is eye contact is present is within this interval.

1b)
```{r}
cat("Confidence interval for Y.2. is [", Y.2. - ME1a, "," , Y.2. + ME1a,"]." )
```
The interpretation of this is that I am 99% confident that the true mean of the treatments where the personnel officer is female is within this interval.

1c)
```{r}
D1 = Y2.. - Y1..
SD = sqrt((2*MSE)/ (b * n))
#bonferroni number
ME1c =qt(1-(0.95/4),(n*b-1)*(n-1)) * SD
cat("Confidence interval for D1 is [", D1 - ME1c, "," , D1 + ME1c,"]." )
cat("Confidence interval for D2 is [", D2 - ME1c, "," , D2 + ME1c,"]." )
```
The true difference of means displayed in these intervals are of the presence of eye contact and the gender of the office personnel.

1d)
```{r}
cat("Confidence interval for Y21.. is [", Y21. - ME1a, "," , Y21. + ME1a,"]." )
```
The interpretation of this is that I am 99% confident that the true mean of the treatments where the personnel officer is male and the eye contact is absent is within this interval.


2a)
