---
title: "Discussion 8 - Ishita Dutta"
output: pdf_document
---


# Problem 2

# (a)
  Using a random sample size of 15 (without replacement) from <exe.pop>, find the 95% confidence interval for mu, and check if the true mean was within the confidence interval.  You should return a TRUE or FALSE (or a 1 or 0).
  
```{r}
mu = 5
exp.pop = rexp(1000000,1/mu)
```

```{r}
sample_2 = sample(exp.pop,15,replace = FALSE)
s1 = t.test(x=sample_2,mu=5,conf.level = 0.95)
s1$conf.int

mu > s1$conf.int[1] & mu < s1$conf.int[2]
```

# (b)
  Repeat  (a)  10000  times,  so  that  you  should  have  a  vector  of  10000  TRUE’s  or FALSE’s depending on if the true mean was within in confidence interval or not.Use the function <table()> on your vector and display the results.
```{r}
repetitions = 10000
store_results = c()
for(i in 1:repetitions){
  sample_1 = sample(exp.pop,15,replace = FALSE)
  s1 = t.test(x=sample_1,mu=5,conf.level = 0.95)
  store_results[i] = mu > s1$conf.int[1] & mu < s1$conf.int[2]
}

table(store_results)
```

# (c)
  Find the coverage probability based on (b).

```{r}
mean(store_results)
```

# (d)
  Did this confidence interval have at least a 95% coverage probability?  Explain why it did, or why it may not have.

__Solution:__ This interval does not have a 95% coverage probability because our distribution is not a normal distribution, it is an exponential distribution, which is not symmetric so we cannot assume a 95% coverage probability. The coverage probability is 0.9114.